---
title: "Strategy to Launch Historical Funding"
---

```{r global_options, include=FALSE}
library(knitr)
library(tidyverse)
library(colorspace)
library(cowplot)
library(dplyr)
library(scales)

opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```


Below is the historical performance of the Strategy-to-Launch life cycle in the BAT.

Comparing the data from STL reporting and DPG reporting.


```{r message = FALSE, echo=FALSE, warning=FALSE}
# load data
STL_spend <- read_csv("data\\STL_funding_levels_no_formatting.csv"
  # , col_types = cols("Roadmap amounts (000's)" = col_double())
  )
STL_spend
# unique(STL_spend$"Roadmap_amounts")
# colnames(STL_spend)
```


Table 1. Raw data


---

```{r message = FALSE, echo=FALSE, warning=FALSE, results=FALSE, fig.width=12}
colnames(STL_spend)

# scaleFactor <- max(STL_spend$Roadmap_projects) / max(STL_spend$Roadmap_amounts)

# plot number of projects and cost by FY
ggplot(STL_spend, aes(x = FY, y = Roadmap_projects, label = Roadmap_amounts)) +
  # set up project visual
    # change scale of data values to align with fund funding scale
    # add vertical bars for funding levels
  geom_col(aes(y = Roadmap_amounts), fill = ifelse(STL_spend$FY=="FY25", "grey", "steelblue"), alpha = 0.8) + 
  geom_line(aes(
    y = Roadmap_projects * 500, group = 1, 
    linetype = "# of projects"), 
    color = "blue", linewidth = 1.7, alpha = 0.7
    ) +
  # add points for project numbers, possibly labels - depends on clutter
    # change scale of points to align with fund funding scale
  geom_point(aes(y = Roadmap_projects * 500), size = 2, color = "blue") +
  scale_y_continuous(
    name = "$ (000's)", 
    # re-scale secondary axis to accommodate previous scalings
    sec.axis=sec_axis(~./500, name="# projects")
    ) +
  theme_bw() + 
  ggtitle("STL Roadmap Funded") + 
  geom_text(
    # use labeler package to represent amounts in millions
    aes(label = label_number_si()(Roadmap_amounts*1000)), 
    vjust = -0.5, 
    color = "white"
    ) +
  theme(
    # simplify background elements
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(), 
    panel.border = element_blank(), 
    plot.title = element_text((hjust = 10)),
    # axis.title.x = element_text(margin = margin(t = 5)),
    # nudge axis text positions
    axis.text.x.bottom = element_text(margin = margin(t = 5)),
    axis.title.y.right = element_text(vjust = 2),
    axis.title.y.left = element_text(vjust = 2),
    legend.position = c(0.95,0.95),
    legend.title = element_blank(),
    ) +
  xlab("Fiscal Year")

# next do for self-funded

```


Figure 1. Roadmap funded projects and funding levels each fiscal year.

---

```{r message = FALSE, echo=FALSE, warning=FALSE, results=FALSE, fig.width=13}
# self-fund
colnames(STL_spend)

# plot number of projects and cost by FY
ggplot(STL_spend, aes(x = FY, y = Roadmap_projects, label = Roadmap_amounts)) +
  # set up project visual
    # change scale of data values to align with fund funding scale
  geom_col(aes(y = Self_amounts), fill = ifelse(STL_spend$FY=="FY25", "grey", "steelblue"), alpha = 0.8) + 
  geom_line(aes(
    y = Self_projects * 100, group = 1, 
    linetype = "# of projects"), 
    color = "blue", linewidth = 1.7, alpha = 0.7
    ) +
  # add points for project numbers, possibly labels - depends on clutter
    # change scale of points to align with fund funding scale
  geom_point(aes(y = Self_projects * 100), size = 2, color = "blue") +
  # add vertical bars for funding levels
  scale_y_continuous(
    name = "$ (000's)", 
    # re-scale secondary axis to accommodate previous scalings
    sec.axis=sec_axis(~./100, name="# projects")
    ) +
  theme_bw() + 
  ggtitle("STL Self-Funded") + 
  geom_text(
    # use labeler package to represent amounts in millions
    aes(label = label_number_si()(Self_amounts*1000)), 
    vjust = -0.5, 
    color = "white"
    ) +
  theme(
    # simplify background elements
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(), 
    panel.border = element_blank(), 
    plot.title = element_text((hjust = 10)),
    # axis.title.x = element_text(margin = margin(t = 5)),
    # nudge axis text positions
    axis.text.x.bottom = element_text(margin = margin(t = 5)),
    axis.title.y.right = element_text(vjust = 2),
    axis.title.y.left = element_text(vjust = 2),
    legend.position = c(0.95,0.95),
    legend.title = element_blank(),
    ) +
  xlab("Fiscal Year")
  # ylim(0, 2000)

```

Figure 2. Self-funded projects and cost each fiscal year.

```{r message = FALSE, echo=FALSE, warning=FALSE, results=FALSE, fig.width=10}
# total by year 
colnames(STL_spend)
ggplot(STL_spend, aes(x = FY, y = Total_projects, group = interaction(FY, Total), label = Total)) +
  geom_col(aes(y = Total), fill = ifelse(STL_spend$FY=="FY25", 'grey', "green4")) + 
  geom_line(aes(
    y = Total_projects * 400,
    group = 1,
    linetype = "# of projects"), 
    color = "green3",
    linewidth = 1.7
  ) +
  theme_bw() +
  ggtitle("Total STL Historical funding") +
    theme(plot.title = element_text(hjust = .44, face = "bold")) +
  # facet_wrap(
  #   vars(FY, Total),
  #   nrow = 4
  # ) +
  geom_point(aes(y = Total_projects * 400),
  color = "green4"
  ) +
  # scale_y_continuous(
  #   name = "Total funding (000's)"
  # ) +
  theme(
    # simplify background elements
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(), 
    panel.border = element_blank(), 
    plot.title = element_text((hjust = 10)),
    # axis.title.x = element_text(margin = margin(t = 5)),
    # nudge axis text positions
    axis.text.x.bottom = element_text(margin = margin(t = 5)),
    axis.title.y.right = element_text(vjust = 2),
    axis.title.y.left = element_text(vjust = 2),
    legend.position = c(0.9,0.9),
    legend.title = element_blank()
    ) +
  xlab("Fiscal Year") +
    geom_text(
    # use labeler package to represent amounts in millions
    aes(label = label_number_si()(Total*1000)), 
    vjust = -0.5, 
    color = "white"
    ) +
  scale_y_continuous(
    name = "Total Funding (000's)", 
    # re-scale secondary axis to accommodate previous scalings
    sec.axis=sec_axis(~./400, name="# projects")
    ) 

```

Fig 3.  STL project totals by fiscal year.

---

The differences in the datasets are not too disparate; attributing
most of the differences to snapshots in time; whether before or after a key decision, 
trends are the same. Using the DPG data will result in have less friction in 
future discussions. Depicted as such.

Funding shifted in FY22 the same time BAT funding definitions changed. Where STL
funding levels were ~10%+ of total BAT initiatives, recent years has STL funding 
at 1-2% of total BAT funding levels. Alos, latter-year trends are grouping more 
projects into large transform efforts, under the new fiscal definitions. 
FY22 and earlier BAT funding categories had double the granularity. 
No correlation exists between project counts and funding levels.

Funding for certain IT systems, like Windchill, has also shifted.  Dell Digital
does not provide support as it did in 2018 and 2020 when STL funded those upgrades.

---